---
title: "p8105_hw1_coo2124"
output: github_document
---

## PROBLEM 1

Install and Load Package

```{r}
data("penguins", package = "palmerpenguins")
```


Inline R Code to Describe the Dataset

The `penguins` dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns, representing data on penguins. The important variables include `species` (the species of the penguin), `island` (the island where the penguin was observed), `bill_length_mm` (the length of the penguin's bill), `flipper_length_mm` (the length of the penguin's flipper), and `body_mass_g` (the body mass of the penguin in grams).

The average flipper length across all penguins is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm.


## the data in this dataset, including names / values of important variables
```{r}
str(penguins)
```

## Create a scatterplot of flipper length vs bill length:

```{r}
library(ggplot2)

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(
    title = "Scatterplot of Flipper Length vs Bill Length",
    x = "Bill Length (mm)",
    y = "Flipper Length (mm)",
    color = "Species"
  ) +
  theme_minimal()
```

## Export scatterplot using ggsave

```{r}
ggsave("flipper_vs_bill_scatterplot.png")
```


## PROBLEM 2

Create a data frame

```{r}
library(tidyverse)

set.seed(123)
df <- tibble(
  normal_sample = rnorm(10),                          
  logical_vector = rnorm(10) > 0,                      
  character_vector = rep(c("a", "b", "c", "d"), length.out = 10),  
  factor_vector = factor(rep(c("level1", "level2", "level3"), length.out = 10))  
)

df
```

 mean of each variable in the dataframe
 
```{r}
mean(df %>% pull(normal_sample))        
mean(df %>% pull(logical_vector))        
mean(df %>% pull(character_vector))     
mean(df %>% pull(factor_vector))         
```
 
 ## What works and what doesn’t?
 
The mean of normal_sample  worked because it's a numeric variable.
The mean of logical_vector  worked because R coerces TRUE to 1 and FALSE to 0, allowing for the calculation of a numeric mean.
The mean of character_vector did not work because the mean of a character vector can not be taken.
The mean of factor_vector did not work because factors are categorical variables, not numeric.


## code chunk that applies the as.numeric function to the logical, character, and factor variables
```{r}
library(dplyr)

as.numeric(df %>% pull(logical_vector))
as.numeric(df %>% pull(character_vector))
as.numeric(df %>% pull(factor_vector))
```

## what happens when you try to take the mean?

Logical variables converted to numeric, where TRUE becomes 1 and FALSE becomes 0.
Character values cannot be directly converted to numbers, resulting in NA.
Factors are internally stored as integers corresponding to their levels, which is why they can be converted to numeric values.
